---
title: "Project: Retail Sale Prediction"
author: "Zach Zhu"
date: "12/04/2019"
header-includes:
- \usepackage{graphicx}
- \usepackage{fancyhdr}
- \usepackage{enumitem}
- \usepackage{amsmath ,amssymb ,amsthm}
- \pagestyle{fancy}
- \rhead{\today}
- \lhead{MSiA401}
- \cfoot{\thepage}
- \renewcommand{\headrulewidth}{1pt}
output: 
  pdf_document:
    latex_engine: xelatex
    toc: false
    toc_depth: 3
    number_sections: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file contains the code to clean the raw data in the file `sales_data.csv` and output data for modelling `cleaning_records.csv`

## Loading dataset and import libraries

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(lubridate)
library(ggplot2) #Plots
```

```{r, message=FALSE, warning=FALSE}
records <- read.csv("sales_data.csv", stringsAsFactors = FALSE)
```


## Standardizing data

### Reformat date variables

```{r}
records <- mutate(records, datead6 = mdy(datead6), datelp6 = mdy(datelp6))
```

### Checking if datead6 <= datelp6

```{r}
test <- records[records$datead6 > records$datelp6,][c('datead6','datelp6')] #834 rows
#For records with "date added" after "date last purchase" - imputing datead6 with datelp6
records <- transform(records, datead6 = if_else(datead6 > datelp6, datelp6, datead6))
```

### Add the first 3 digits into `lpuryear`

```{r}
records[!is.na(records$lpuryear),]$lpuryear <- ifelse(records[!is.na(records$lpuryear),]$lpuryear<3,as.integer(paste0('201',as.character(records[!is.na(records$lpuryear),]$lpuryear))),as.integer(paste0('200',as.character(records[!is.na(records$lpuryear),]$lpuryear))))
```

### Add customer id

Adding an ID column to identify each customer uniquely

```{r}
records <- mutate(records, custId = as.integer(row.names(records)))
```


## Handling missing and incorrect data

## Summary Statistics of columns
```{r}
summary(records)
#Fields with missing values
#1. lpuryear
colnames(records)
cor(records[c("targdol","slstyr","slslyr","sls2ago","sls3ago","slshist","ordtyr","ordlyr",
              "ord2ago","ord3ago","ordhist","falord","sprord")])
```

### datelp6 and lpuryear

We found some NA values in `lpuryear`. To deal with NA values, we use the year in `datelp6` to fill it. We also compare `lpuryear` with `datelp6` and update one of them by the larger one. 

```{r}
records[is.na(records$lpuryear),]$lpuryear <- year(records[is.na(records$lpuryear),]$datelp6)

test <- records[records$lpuryear < year(records$datelp6),] #3553 cases
test_1 <- records[records$lpuryear > year(records$datelp6),] #4219 cases

#Imputing lpuryear with max of the two years 
records$lpuryear <- with(records, pmax(lpuryear, year(datelp6)))
#records[records$lpuryear < year(records$datelp6),]$lpuryear <- year(records[records$lpuryear < records$datelp6,]$datelp6)
```

## Compare sales($) fields and order(#) fields
```{r}
test <- records[(records$slstyr > 0 && records$ordtyr == 0),] #0
test <- records[(records$slslyr > 0 && records$ordlyr == 0),] #0
test <- records[(records$sls2ago > 0 && records$ord2ago == 0),] #0
test <- records[(records$sls3ago > 0 && records$ord3ago == 0),] #0
test <- records[(records$slshist > 0 && records$ordhist == 0),] #0

temp <- filter(records, slstyr == 0 & (ordtyr > 0)) #205
temp <- filter(records, slslyr == 0 & (ordlyr > 0)) #121
temp <- filter(records, sls2ago == 0 & (ord2ago > 0)) #124
temp <- filter(records, sls3ago == 0 & (ord3ago > 0)) #237
temp <- filter(records, slshist == 0 & (ordhist > 0)) #499

test <- records[(records$slstyr < records$ordtyr),] #206
#Imputing ordtyr with slstyr, i.e. 0
records$ordtyr <- if_else((records$slstyr < records$ordtyr),records$slstyr,records$ordtyr)

test <- records[(records$slslyr < records$ordlyr),] #121
#Imputing ordlyr with slslyr, i.e. 0
records$ordlyr <- if_else((records$slslyr < records$ordlyr),records$slslyr,records$ordlyr)

test <- records[(records$sls2ago < records$ord2ago),] #124
#Imputing ord2ago with sls2ago, i.e. 0
records$ord2ago <- if_else((records$sls2ago < records$ord2ago),records$sls2ago,records$ord2ago)

test <- records[(records$sls3ago < records$ord3ago),] #237
#Imputing ord3ago with sls3ago, i.e. 0
records$ord3ago <- if_else((records$sls3ago < records$ord3ago),records$sls3ago,records$ord3ago)

test <- records[(records$slshist < records$ordhist),] #500
#Imputing ordhist with slshist, i.e. 0
records$ordhist <- if_else((records$slshist < records$ordhist),records$slshist,records$ordhist)

```

### Handling incorrect `falord`, `sprord` and `ordhist`

If the sum of `falord` and `sprord` is greater than `ordhist`, we need to update `ordhist`; otherwise, we need to update either `falord` or `sprord`.

If the sum of `falord` and `sprord` are greater than `ordhist`, update `ordhist`. Otherwise, update either `falord` or `sprord` with respect to `lpurseason`,
```{r}
records %>%
    mutate(ordhist = ifelse(ordhist<falord+sprord, falord+sprord, ordhist)) %>%
    mutate(sprord = ifelse(ordhist>falord+sprord & month(records$datelp6) < 7,
                           ordhist-falord, sprord)) %>%
    mutate(falord = ifelse(ordhist>falord+sprord & month(records$datelp6) >= 7,
                           ordhist-sprord, falord))
```

## Handling incorrect unmatched order number and sales for each year

```{r}
test <- records[records$datead6 > records$datelp6,][c('datead6','datelp6')] #834 rows
#For records with "date added" after "date last purchase" - imputing datead6 with datelp6
records <- transform(records, datead6 = if_else(datead6 > datelp6, datelp6, datead6))

```

## Creating binary target to identify responders for logistic regression model
```{r}
records$responder <- ifelse((records$targdol > 0), 1, 0)

```

## Writing out clean data

```{r}
write.csv(train, file = "cleaning_records.csv", row.names=FALSE)
```